---
title: "Assignment2_Part1_VoiceInSchizophrenia"
author: "Amalie Lysgaard Andersen"
date: "July 17, 2016"
output: html_document
---

```{r setup, include=FALSE}
setwd("~/Library/Mobile Documents/com~apple~CloudDocs/4. Aarhus Universitet/3. Semester/1. Experimental Methods lll/R FOLDER/Schizophrenia_updated/Schizophrenia_updated")
locpath = "~/Library/Mobile Documents/com~apple~CloudDocs/4. Aarhus Universitet/3. Semester/1. Experimental Methods lll/R FOLDER/Schizophrenia_updated/Schizophrenia_updated"

#setwd("~/Library/Mobile Documents/com~apple~CloudDocs/Aarhus Universitet/3. Semester/1. Experimental Methods lll/R FOLDER/Schizophrenia ")
#locpath = "~/Library/Mobile Documents/com~apple~CloudDocs/Aarhus Universitet/3. Semester/1. Experimental Methods lll/R FOLDER/Schizophrenia"

library(pacman)
p_load(lmerTest, pastecs, ggplot2, tidyverse, gdata, MuMIn, effects, stringr, plyr, Metrics, vtreat, cvTools, simr, lme4, stringi, stringr)

#update.packages("lmer")
#update.packages("stringr")

#install.packages("stringr")

```

## Assignment 2 - Part 1 - Assessing voice in schizophrenia

Schizophrenia has been associated with "inappropriate" voice, sometimes monotone, sometimes croaky. A few studies indicate that pitch might be an index of schizophrenia. However, an ongoing meta-analysis of the literature (which you will have a go at in the last assignment) indicates that pitch mean and standard deviation are only weak indicators of diagnosis. Can we do better with our new fancy complex skills?

The corpus you are asked to analyse is a set of voice recordings from people with schizophrenia (just after first diagnosis) and 1-1 matched controls (on gender, age, education). Each participant watched 10 videos of triangles moving across the screen and had to describe them (so you have circa 10 recordings per person). We have already extracted the pitch once every 10 milliseconds and you will have to use this data to assess differences in the voice.

Can you characterize voice in schizophrenia as acoustically different? Report the methods you used to answer this question and the results from the analyses. Add a couple of lines trying to interpret the results (make sense of the difference). E.g. People with schizophrenia tend to have high-pitched voice, and present bigger swings in their prosody than controls. Add a couple of lines describing limitations of the data/analyses if any is relevant.


N.B. There are looots of files to be dealt with. Maybe too many for your computer, depending on how you load the files. This is a challenge for you. Some (complementary) possible strategies:
- You can select a subset of files only (and you have to justify your choice).
- You can learn how to use the apply() or map() functions.
- You can coordinate with classmates.

Hint: There is some information in the filenames that you might need.
Hint: It might be a good idea to first create a function which loads and parses one file, and then loop through the list of files to read them all. For example

```{r + Simply one file + INCLUDE = F}
#Function to extract descriptors for one file
pitch.data = function(x)
  {
  result<-c(
  mean=mean(x$f0, na.rm=TRUE),
  median=median(x$f0, na.rm=TRUE),
  max.min=range(x$f0, na.rm=TRUE),
  range=max(x$f0, na.rm=TRUE) - min(x$f0, na.rm=TRUE),
  variance=var(x$f0, na.rm=TRUE),
  std.Dev=sd(x$f0, na.rm=TRUE),
  iqr=IQR(x$f0, na.rm = TRUE),
  mad=mad(x$f0, na.rm = TRUE),
  coefVar=sd(x$f0, na.rm=TRUE)/mean(x$f0, na.rm=TRUE)*100,
  se=sd(x$f0, na.rm=TRUE)/sqrt(length(x$f0[!is.na(x)])),
  quantile=quantile(x$f0, na.rm=TRUE)
  )
return(result)
}

p <- read.table("Study1D0S101T1_f0.txt", sep="\t", header = TRUE)
pitch.data(p)
```

```{r Jan's}
read_pitch <- function(filename) {
    read.delim("Study1D0S101T1_f0.txt")
    # parse filename
    # extract descriptors
    # combine all this data
    return(data)
  }

# test it on just one file while writing the function
#test_data = read_pitch("Pitch/test_file")

### JAN
# when you've created a function that works, you can
read_pitch <- function(filename) {
    raw <- read.delim(filename)
    name <- as.character(filename)
    
    data <- cbind(raw, name) %>%
      mutate(nrow = nrow(raw)) %>%
      summarise(mean = mean(f0),
                sd = sd(f0),
                range = max(f0) - min(f0),
                iqr = IQR(f0),
                median = median(f0),
                mean_abs = mad(f0),
                coef_var = (sd / mean)*100,
                nrow = max(nrow),
                se = sd / sqrt(nrow))
    
    data <- data %>%
      mutate(name = str_remove_all(name, "Pitch/"),
         name = str_remove_all(name, ".txt"),
         name = str_remove_all(name, "_f0")) %>%
      
      mutate(study = substr(name, 6, 6),
         diagnosis = substr(name, 8, 8),
         ID = substr(name, 10, 12),
         triangle = substr(name, 14, 15)) %>%
      
      select(-name) %>%
      select(ID, study, diagnosis, triangle, nrow, mean:coef_var, se)
    
    return(data)
    }

# Creating list of files in the folder 'Pitch'
pitch_list <- list.files(path = locpath, recursive = TRUE, pattern = "*f0.txt")

# When you've created a function that works, you can
# path = paste0("Pitch/", list.files("Pitch/"))
pitch_data = pitch_list %>% 
    purrr::map_df(read_pitch)

# Rename
pitch_data <- rename.vars(pitch_data, "triangle", "trial")
pitch_data$diagnosis <- ifelse(pitch_data$diagnosis == "0", "Control", "Schizophrenia")

write_csv(pitch_data, "data.csv")
```

```{r Test}
#Test file
test_data = read_pitch("Study1D0S101T1_f0.txt")
test_data

#We already did it all 
```

```{r Readying data}
# Load data
data <- read.csv("data.csv")

# Loading demographic data
demo <- read.delim("DemoData.txt", header = T)
demo <- rename.vars(demo,"Subject", "ID")
demo <- rename.vars(demo,"Study", "study")
data <- rename.vars(data, "diagnosis", "Diagnosis")

#data$Diagnosis <- ifelse(data$Diagnosis == "control", "Control", "Schizophrenia")

# Merging datafiles
df <- join(data, demo) 
```

1. In the course of this assignment you have to first select one datafile and figure out how to:

- Extract "standard" descriptors of pitch: Mean, standard deviation, range
- Extract less "standard" descriptors of pitch you can think of (e.g. median, iqr, mean absoluted deviation, coefficient of variation)

2. Second you will have to turn the code into a function and loop through all the files (or even better use apply/sapply/lapply)
- Remember to extract the relevant information from the file names (Participant, Diagnosis, Trial, Study)

### Exercise 3. 
Make one model per acoustic feature and test whether you can observe significant difference due to Diagnosis. Tip: Which other fixed factors should you control for (that is, include in the model)? Which random ones?
- Bonus points: cross-validate the model and report the betas and standard errors from all rounds to get an idea of how robust the estimates are. 

```{r Mean model}
## #Mean
stat.desc(df$mean, norm = T)
df$log_mean<-log(df$mean)
stat.desc(df$log_mean, norm = T)
lm.mean <- lmer(log_mean ~ Diagnosis + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(lm.mean)

#Assumptions
plot(residuals(lm.mean))
qqnorm(residuals(lm.mean))
hist(residuals(lm.mean))

```

```{r Median model}
### Median
stat.desc(df$median, norm = T)
df$log_median<-log(df$median)
stat.desc(df$log_median, norm = T)
lm.median <- lmer(log_median ~ Diagnosis + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(lm.median)

#Assumptions
plot(residuals(lm.median))
qqnorm(residuals(lm.median))
hist(residuals(lm.median))

```

```{r SD model}
#Standard Deviation
df$log_sd<-log(df$sd)
stat.desc(df$log_sd, norm = T)
lm.std <- lmer(log_sd ~ Diagnosis + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(lm.std)

#Assumptions
plot(residuals(lm.std))
qqnorm(residuals(lm.std))
hist(residuals(lm.std))

```

```{r Range mdoel}
#Range
df$log_range<-log(df$range)
stat.desc(df$log_range, norm = T)
lm.range <- lmer(log_range ~ Diagnosis + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(lm.range)
#Assumptions
plot(residuals(lm.range))
qqnorm(residuals(lm.range))
hist(residuals(lm.range))
```

```{r MAD model}
#Median Absolute Deviation
df$log_mad<-log(df$mean_abs)
stat.desc(df$log_mad, norm = T)
lm.mad <- lmer(log_mad ~ Diagnosis + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(lm.mad)
#Assumptions
plot(residuals(lm.mad))
qqnorm(residuals(lm.mad))
hist(residuals(lm.mad))

```

```{r CofVar model}
#Coefficient of Variation
df$log_cv<-log(df$coef_var)
stat.desc(df$log_cv, norm = T)
lm.cv <- lmer(log_cv ~ Diagnosis + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(lm.cv)
#Assumptions
plot(residuals(lm.cv))
qqnorm(residuals(lm.cv))
hist(residuals(lm.cv))

```

```{r IQR model}
#Interquartile range
df$log_iqr<-log(df$iqr)
stat.desc(df$log_iqr, norm = T)
lm.iqr <- lmer(log_iqr ~ Diagnosis + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(lm.iqr)
#Assumptions
plot(residuals(lm.iqr))
qqnorm(residuals(lm.iqr))
hist(residuals(lm.iqr))

```

```{r SE model}
#Standard Error
df$log_se<-log(df$se)
stat.desc(df$log_se, norm = T)
lm.se <- lmer(log_se ~ Diagnosis + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(lm.se)
#Assumptions
plot(residuals(lm.se))
qqnorm(residuals(lm.se))
hist(residuals(lm.se))
```


- Bonus points: cross-validate the model and report the betas and standard errors from all rounds to get an idea of how robust the estimates are.

```{r CV of models}
p_load(caret, Metrics)

df <- na.omit(df)
folds = createFolds(unique(df$ID), 5)
folds<-lapply(folds, function(x) unique(df$ID)[x]) #to rename them what they are actually called in the data
folds

df$model1_pred<- 0
df$model2_pred<- 0
df$model3_pred<- 0
df$model4_pred<- 0
df$model5_pred<- 0
df$model6_pred<- 0
df$model7_pred<- 0
df$model8_pred<- 0

for (f in folds) {
  idx = df$ID %in% f
  train = filter(df, !idx)
  test = filter(df, idx)
  
  model1 = lmer(log_mean ~ Diagnosis + Gender + (1 | ID) + (1 | trial), train, REML = F)
  model2 = lmer(log_median ~ Diagnosis + Gender + (1 | ID) + (1 | trial), train, REML = F)
  model3 = lmer(log_range ~ Diagnosis + Gender + (1 | ID) + (1 | trial), train, REML = F)
  model4 = lmer(log_sd ~ Diagnosis + Gender + (1 | ID) + (1 | trial), train, REML = F)
  model5 = lmer(log_mad ~ Diagnosis + Gender + (1 | ID) + (1 | trial), train, REML = F)
  model6 = lmer(log_cv ~ Diagnosis + Gender + (1 | ID) + (1 | trial), train, REML = F)
  model7 = lmer(log_iqr ~ Diagnosis + Gender + (1 | ID) + (1 | trial), train, REML = F)
  model8 = lmer(log_se ~ Diagnosis + Gender + (1 | ID) + (1 | trial), train, REML = F)
  
  df$model1_pred[idx] = predict(model1, newdata = test, allow.new.levels = TRUE)
  df$model2_pred[idx] = predict(model2, newdata = test, allow.new.levels = TRUE)
  df$model3_pred[idx] = predict(model3, newdata = test, allow.new.levels = TRUE)
  df$model4_pred[idx] = predict(model4, newdata = test, allow.new.levels = TRUE)
  df$model5_pred[idx] = predict(model5, newdata = test, allow.new.levels = TRUE)
  df$model6_pred[idx] = predict(model6, newdata = test, allow.new.levels = TRUE)
  df$model7_pred[idx] = predict(model7, newdata = test, allow.new.levels = TRUE)
  df$model8_pred[idx] = predict(model8, newdata = test, allow.new.levels = TRUE)
}

rmse(df$log_mean, df$model1_pred)
# [1] 0.233166
rmse(df$log_median, df$model2_pred)
# [1] 0.2579872
rmse(df$log_range, df$model3_pred)
# [1] 0.4793548
rmse(df$log_sd, df$model4_pred)
# [1] 0.6196506
rmse(df$log_mad, df$model5_pred)
# [1] 0.6095942
rmse(df$log_cv, df$model6_pred)
# [1] 0.6363469
rmse(df$log_iqr, df$model7_pred)
# [1] 0.7760934
rmse(df$log_se, df$model8_pred)
# [1] 0.7172258
```


### 3a. Is study a significant predictor in these models? What should you infer from this? Does study interact with diagnosis? What should you infer from this?
```{r Models}
#Mean
st.mean <- lmer(log_mean ~ Diagnosis + study + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(st.mean)

#Interaction
in.mean <- lmer(log_mean ~ Diagnosis*study + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(in.mean)

anova(st.mean, in.mean)




#Median
st.median <- lmer(log_median ~ Diagnosis + study + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(st.median)

#Interaction
in.median <- lmer(log_median ~ Diagnosis*study + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(in.median)

anova(st.median, in.median)




#Standard Deviation
st.std <- lmer(log_sd ~ Diagnosis + study + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(st.std)

#Interaction
in.std <- lmer(log_sd ~ Diagnosis*study + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(in.std)

anova(st.std, in.std)




#Range
st.range <- lmer(log_range ~ Diagnosis + study + Gender+ (1 | ID) + (1 | trial), df, REML = F)
summary(st.range)

#Interaction
in.range <- lmer(log_range ~ Diagnosis*study + Gender+ (1 | ID) + (1 | trial), df, REML = F)
summary(in.range)

anova(st.range, in.range)




#Meadian Absolute Deviation
st.mad <- lmer(log_mad ~ Diagnosis + study + Gender+ (1 | ID) + (1 | trial), df, REML = F)
summary(st.mad)

#Interaction
in.mad <- lmer(log_mad ~ Diagnosis*study + Gender+ (1 | ID) + (1 | trial), df, REML = F)
summary(in.mad)

anova(st.mad, in.mad)
#Study not significant



#Coefficient of Variation
st.cv <- lmer(log_cv ~ Diagnosis + study + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(st.cv)

#Interaction
in.cv <- lmer(log_cv ~ Diagnosis*study + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(in.cv)

anova(st.cv, in.cv)
#Study not significant




#Interquartile range
st.iqr <- lmer(log_iqr ~ Diagnosis + study + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(st.iqr)

#Interaction
in.iqr <- lmer(log_iqr ~ Diagnosis*study + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(in.iqr)

anova(st.iqr, in.iqr)
#Study not significant




#Standard Error
st.se <- lmer(log_se ~ Diagnosis + study + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(st.se)

#Interaction
in.se <- lmer(log_se ~ Diagnosis*study + Gender + (1 | ID) + (1 | trial), df, REML = F)
summary(in.se)

anova(st.se, in.se)

```

```{r Plots}
p_load(effects)


ggplot(df,
aes(mean, Diagnosis,colour=Gender)) +
geom_point() +
theme_classic()
#Outliers

df$Diagnosis <- as.factor(df$Diagnosis)


plot(allEffects(in.mean))
plot(allEffects(in.median))
plot(allEffects(in.range))
plot(allEffects(in.std))

plot(allEffects(in.mad))
plot(allEffects(in.iqr))
plot(allEffects(in.cv))
plot(allEffects(in.se))
```




















### CODE DUMP

```{r Our load loop}
#Preparing columns 
id = NULL
study = NULL
diagnosis = NULL
trial = NULL
std = NULL
var = NULL
mean = NULL
range = NULL
median = NULL
iqr = NULL
mad = NULL
x = NULL
n = 1


# The loop 
for (i in pitch_list) {
  x = read.delim(i, header = T)
  x = x$f0
  id[n] = str_extract(i, "S+\\d+")
  id = str_extract(id,"\\d+")
  diagnosis[n] = str_extract(str_extract(i,"D+\\d"), "\\d")
  trial[n] = str_extract(str_extract(i,"T.+\\d"), "\\d+")
  study[n] = str_extract(i, "\\d")
  range[n] = range(x,na.rm = T)
  mean[n] = mean(x,na.rm = T)
  var[n] = var(x, na.rm=TRUE)
  std[n] = sd(x)
  median[n] = median(x, na.rm = T)
  iqr[n] = IQR(x,na.rm = T)
  mad[n] = mad(x,na.rm = T)
  n = n+1
}

dframe = data.frame(id,study,diagnosis,trial,range,mean,var,std,median,mad, iqr)
```

```{r Our loop: From nanna}
  
#Creating list of files in the folder 'Pitch'
pitch_list<- list.files(path = locpath, recursive = TRUE, pattern = "*f0.txt")

#Preparing columns
id = NULL
study = NULL
diagnosis = NULL
trial = NULL
std = NULL
var = NULL
mean = NULL
range = NULL
median = NULL
iqr = NULL
mad = NULL
x = NULL
n = 1
 

#The loop 

for (i in pitch_list) {
  x = read.delim(i, header = T)
  x = x$f0
  id[n] = str_extract(i, "S+\\d+")
  id = str_extract(id,"\\d+")
  diagnosis[n] = str_extract(str_extract(i,"D+\\d"), "\\d")
  trial[n] = str_extract(str_extract(i,"T.+\\d"), "\\d+")
  study[n] = str_extract(i, "\\d")
  range[n] = max(x) - min(x)
  mean[n] = mean(x,na.rm = T)
  var[n] = var(x, na.rm=TRUE)
  std[n] = sd(x)
  median[n] = median(x, na.rm = T)
  iqr[n] = IQR(x,na.rm = T)
  mad[n] = mad(x,na.rm = T)
  n = n+1
}

dframe = data.frame(id,study,diagnosis,trial,range,mean,var,std,median,mad,iqr)

data <-  dframe
```

```{r}
 
#Looping but not really

pitch_data = pitch_list %>%
    purrr::map_df(read_pitch)

write_csv(pitch_data, "data.csv")

data <- read.csv("data.csv")

data$diagnosis <- ifelse(data$diagnosis == "0", "Control", "Schizophrenia")

```

```{r}
demodata <- read.table("Assignment3_VoiceSchizo_DemoData.txt", header = T)

demodata <- rename.vars(demodata, "Subject", "ID")
demodata <- rename.vars(demodata, "Study", "study")
demodata <- rename.vars(demodata, "Diagnosis", "diagnosis")

df <- join(data, demodata)

write.csv(df, "data_with_demo.csv")

```

```{r + Simply one file + INCLUDE = F}
#Function to extract descriptors for one file
pitch.data = function(x)
  {
  result<-c(
  mean=mean(x$f0, na.rm=TRUE),
  median=median(x$f0, na.rm=TRUE),
  max.min=range(x$f0, na.rm=TRUE),
  range=max(x$f0, na.rm=TRUE) - min(x$f0, na.rm=TRUE),
  variance=var(x$f0, na.rm=TRUE),
  std.Dev=sd(x$f0, na.rm=TRUE),
  iqr=IQR(x$f0, na.rm = TRUE),
  mad=mad(x$f0, na.rm = TRUE),
  coefVar=sd(x$f0, na.rm=TRUE)/mean(x$f0, na.rm=TRUE)*100,
  se=sd(x$f0, na.rm=TRUE)/sqrt(length(x$f0[!is.na(x)])),
  quantile=quantile(x$f0, na.rm=TRUE)
  )
return(result)
}

p <- read.table("Study1D0S101T1_f0.txt", sep="\t", header = TRUE)
pitch.data(p)
```

```{r Jan's}
read_pitch <- function(filename) {
    read.delim("Study1D0S101T1_f0.txt")
    # parse filename
    # extract descriptors
    # combine all this data
    return(data)
  }

# test it on just one file while writing the function
#test_data = read_pitch("Pitch/test_file")

### JAN
# when you've created a function that works, you can
read_pitch <- function(filename) {
    raw <- read.delim(filename)
    name <- as.character(filename)
    
    data <- cbind(raw, name) %>%
      mutate(nrow = nrow(raw)) %>%
      summarise(mean = mean(f0),
                sd = sd(f0),
                range = max(f0) - min(f0),
                iqr = IQR(f0),
                median = median(f0),
                mean_abs = mad(f0),
                coef_var = (sd / mean)*100,
                nrow = max(nrow),
                se = sd / sqrt(nrow))
    
    data <- data %>%
      mutate(name = str_remove_all(name, "Pitch/"),
         name = str_remove_all(name, ".txt"),
         name = str_remove_all(name, "_f0")) %>%
      
      mutate(study = substr(name, 6, 6),
         diagnosis = substr(name, 8, 8),
         ID = substr(name, 10, 12),
         triangle = substr(name, 14, 15)) %>%
      
      select(-name) %>%
      select(ID, study, diagnosis, triangle, nrow, mean:coef_var, se)
    
    return(data)
    }

# Creating list of files in the folder 'Pitch'
pitch_list <- list.files(path = locpath, recursive = TRUE, pattern = "*f0.txt")

# When you've created a function that works, you can
# path = paste0("Pitch/", list.files("Pitch/"))
pitch_data = pitch_list %>% 
    purrr::map_df(read_pitch)

# Rename
pitch_data <- rename.vars(pitch_data, "triangle", "trial")
pitch_data$diagnosis <- ifelse(pitch_data$diagnosis == "0", "Control", "Schizophrenia")

write_csv(pitch_data, "data.csv")

```

